Fully structured proof theory \\for intuitionistic modal logics \\(working title)

### 2020-06-01 bicolor NS/aiml20/aiml20.tex
working on aiml abstract

  \begin{abstract}
  We present a nested sequent system for intuitionistic modal logic in which we capture explicitly two relations: one for the accessibility relation associated with the Kripke semantics for normal modal logics (represented with one bracket in a formula or context) and one for the preorder relation associated with the Kripke semantics for intuitionistic logic (represented with two brackets in a formula or context).
  %
  \end{abstract}

  \begin{keyword}
  Nested sequents, Intuitionistic modal logic, Proof theory.
  \end{keyword}
 \end{frontmatter}

@@

\section{Introduction}
\todo{Say: working on multi-conclusion nested sequents for intuitionistic modal logic. A bit of story? Why is useful?}

Structural proof theoretic accounts of intuitionistic modal logic can adopt the paradigm of \emph{labelled deduction} in the form of labelled natural deduction and sequent systems \cite{simpson1994proof},
or the one of \emph{unlabelled deduction} in the form of nested sequent systems \cite{strassburger2013cut,}.

\section{Intuitionistic modal logics}
The language of {intuitionisitic modal logic} is the one of intuitionistic propositional logic with the modal operators $\BOX$ and $\DIA$, standing most generally for \emph{necessity} and \emph{possibility}.
%
Starting with a set $\mathcal{A}$ of atomic propositions, denoted $a$, modal formulas are constructed from the grammar:
%
$$
A \coloncolonequals
a \mid A \AND A \mid \TOP \mid A \OR A \mid \BOT \mid A \IMP A \mid \BOX A \mid \DIA A
$$
%
The axiomatisation that is now generally accepted as intuitionistic modal logic $\IK$ was given by Plotkin and Stirling~\cite{plotkin1986} and is equivalent to the one proposed by Fischer-Servi~\cite{fischer1984}.
%, and by Ewald~\cite{Ewald} in the case of intuitionistic tense logic. 
%
It is obtained from intuitionistic propositional logic by adding:
\begin{itemize}
	\item the \emph{necessitation rule}: $\BOX A$ is a theorem if $A$ is a theorem; and
	\item the following five variants of the \emph{distributivity axiom}:
	\begin{equation*}
	\label{eq:ik}%\hskip-2em
	\begin{array}[t]{r@{\;}l@{\quad}r@{\;}l@{\quad}r@{\;}l}
	\kax[1]\colon&\BOX(A\IMP B)\IMP(\BOX A\IMP\BOX B)
	&
	\kax[3]\colon&\DIA(A\OR B)\IMP(\DIA A\OR\DIA B)
	&
	\kax[5]\colon&\DIA\BOT\IMP\BOT
	\\
	\kax[2]\colon&\BOX(A\IMP B)\IMP(\DIA A\IMP\DIA B)
	&
	\kax[4]\colon&(\DIA A\IMP \BOX B)\IMP\BOX(A\IMP B)\\%x[1ex]
	\end{array}
	\end{equation*}
\end{itemize}

@@

\begin{definition}
	A \emph{bi-relational frame} $\F$ is a triple $\langle W, R, \le \rangle$ 
	%	of a non-empty set of worlds $W$ equipped with two binary relations $R$ and $\le$, where $R$ being the modal \emph{accessibility relation} and $\le$ a preorder (\emph{i.e.} a reflexive and transitive relation), satisfying the following conditions:
	of a set of worlds $W$ equipped with an {accessibility relation} $\rel$ and a preorder $\le$ satisfying:
	\begin{enumerate}
		\item[($F_1$)] For $u, v, v' \in W$, if $u \rel v$ and $v \le v'$, there exists $u'$ s.t.~$u \le u'$ and $u' \rel v'$.

		\item[($F_2$)] For $u', u, v \in W$, if $u \le v$, there exists $v'$ s.t.~$u' \rel v'$ and $v\le v'$.
	\end{enumerate}
	%	
\end{definition}

\marianela{We recall the birelational models \cite{plotkin1986, ewald1986} for intuitionistic modal logics, which are a combination of the Kripke semantics for propositional intuitionistic logic and the one for classical modal logic. A \emph{birelational frame} $\F = \langle W, R, \le \rangle$ is a non-empty set $W$ of worlds together with two binary relations $\le, \rel \subseteq W \times W$, where $\le$ is a preorder(i.e., reflexive and transitive), such that the following two conditions hold:
}

\begin{definition}
	A \emph{bi-relational model} $\M$ is a quadruple $\langle W, R,\le,V \rangle$ with $\langle W, R, \le \rangle$ a bi-relational frame and $V\colon W \to 2^\mathcal{A}$ a monotone valuation function, that is, a function mapping each world $w$ to the subset of propositional atoms true at $w$, additionally subject to:
	if $w \le w'$ then $V(w)\subseteq V(w')$.
\end{definition}

We write $w \Vdash a$ if $a \in V(w)$, and by definition, we always have $w \Vdash \top$ and never that $w \Vdash \bot$. 
%
Then the relation is extended to all formulas by induction, following the rules for both intuitionistic and modal Kripke models:

$w \Vdash A \AND B$ iff $w \Vdash A$ and $w \Vdash B$

$w \Vdash A \OR B$ iff $w \Vdash A$ or $w \Vdash B$

$w \Vdash A \IMP B$ iff for all $w'$ with $w \le w'$, if $w' \Vdash A$ then $w' \Vdash B$

$w \Vdash \BOX A$ iff for all $w'$ and $u$ with $w \le w'$ and $w'Ru$, $u \Vdash A$ %\hfill $(\ast)$

$w \Vdash \DIA A$ iff there exists a $u$ such that $wRu$ and $u \Vdash A$.

\begin{definition}
	A formula $A$ is \emph{satisfied} in a model $\M = \langle W, R, \le, V \rangle$, if for all $w \in W$ we have $w \Vdash A$.
	%
	A formula $A$ is \emph{valid} in a frame $\F = \langle W, R, \le \rangle$, if for all valuations $V$, $A$ is satisfied in $\langle W, R, \le, V \rangle$.
\end{definition}

@@

\begin{theorem}[Fischer-Servi~\cite{Fischer}, Plotkin and Stirling~\cite{Plotkin}]\label{thm:plotkin}
	A formula $A$ is a theorem of $\IK$ if and only if $A$ is valid in every bi-relational frame.
\end{theorem}

@@

\begin{figure}%[h]
	\centering
	\small
	\fbox{
		\begin{minipage}{.95\textwidth}
			\begin{tabular}{@{\!}c@{\quad}c}
				$\vlinf{\rn{id}}{}{\B, \Left, \labels{x}{A} \SEQ \Right, \labels{x}{A} }{}$
				&
				$\vlinf{\llabrn\bot}{}{\B, \Left, \labels{x}{\BOT} \SEQ \Right}{}$
%				\quad
%				$\vlinf{\rlabrn\top}{}{\B, \Left \SEQ \Right, \labels{x}{\TOP}}{}$
				\\\\
				$\vlinf{\llabrn\AND}{}{\B,\Left, \labels{x}{A \AND B} \SEQ \Right}{\B, \Left, \labels{x}{A}, \labels{x}{B} \SEQ \Right}$
				&
				$\vliinf{\rlabrn\AND}{}{\B,\Left \SEQ \Right, \labels{x}{A \AND B}}{\B, \Left \SEQ \Right, \labels{x}{A}}{\B, \Left \SEQ \Right, \labels{x}{B}}$
				\\\\
				$\vliinf{\llabrn\OR}{}{\B, \Left, \labels{x}{A \OR B} \SEQ \Right}{\B, \Left, \labels{x}{A} \SEQ \Right}{\B, \Left, \labels{x}{B} \SEQ \Right}$
				&
				$\vlinf{\rlabrn\OR}{}{\B, \Left \SEQ \Right, \labels{x}{A \OR B}}{\B, \Left \SEQ \Right, \labels{x}{A}, \labels{x}{B}}$
				\\\\
				\multicolumn{2}{c}{
					$\vlinf{\llabrn\IMP}{\text{\scriptsize $y$ fresh}}{\B, \Left \SEQ \Right, \labels{x}{A \IMP B}}{\B, \Left, x \le y, \labels{y}{A} \SEQ \Right, \labels{y}{B}}$
				}
				\\\\
				\multicolumn{2}{c}{
					$\vliinf{\rlabrn\IMP}{}{\B, x \le y, \Left, \labels{x}{A} \SEQ B \SEQ \Right}{\B, x \le y, \Left \SEQ \Right, \labels{y}{A}}{\B, x \le y, \Left, \labels{y}{B} \SEQ \Right}$ 
					% 		&$\vliinf{\rlabrn\IMP}{\text{\scriptsize $x \le y \in \B$}}{\B, \Left, \labels{x}{A \IMP B} \SEQ \Right}{\B, \Left \SEQ \Right, \labels{y}{A}}{\B, \Left, \labels{y}{B} \SEQ \Right}$
				}
				\\\\
				$\vlinf{\llabrn\BOX}{}{\B, \Left, x \le y, yRz, \labels{x}{\BOX A} \SEQ \Right}{\B,\Left, x \le y, yRz, \labels{x}{\BOX A}, \labels{z}{A} \SEQ \Right}$
				&
				$\vlinf{\rlabrn\BOX}{\text{\scriptsize $y, z$ fresh}}{\B, \Left \SEQ \Right, \labels{x}{\BOX A}}{\B, \Left, x \le y, y \rel z \SEQ \Right, \labels{z}{A}}$
				\\\\
				$\vlinf{\llabrn\DIA}{\text{\scriptsize $y$ fresh}}{\B, \Left, \labels{x}{\DIA A} \SEQ \Right}{\B, \Left, x \rel y, \labels{y}{A} \SEQ \Right}$
				&
				$\vlinf{\rlabrn\DIA}{}{\B, \Left, x \rel y \SEQ \Right, \labels{x}{\DIA A}}{\B, \Left, x \rel y \SEQ \Right, \labels{x}{\DIA A}, \labels{y}{A}}$
				\\
				\multicolumn{2}{c}{
					$\mbox{\hbox to .9\linewidth{\dotfill}}$
				}
				\\
				$\vlinf{\rn{refl}}{}{\B, \Left \SEQ \Right}{\B, x\le x, \Left \SEQ \Right}$
				&
				$\vlinf{\rn{trans}}{}{\B, x \le y, y \le z, \Left \SEQ \Right}{\B, x \le y, y \le z, x \le z, \Left \SEQ \Right}$
				\\\\
				\multicolumn{2}{c}{
					$\vlinf{\rn{F_1}}{\text{\scriptsize $u$ fresh}}{\B, xRy, y \le z, \Left \SEQ \Right}{\B, xRy, y \le z, x \le u, uRz, \Left \SEQ \Right}$
				}
				\\\\
				\multicolumn{2}{c}{
					$\vlinf{\rn{F_2}}{\text{\scriptsize $u$ fresh}}{\B, xRy,x \le z, \Left \SEQ \Right}{\B, xRy, x \le z, y \le u, zRu, \Left \SEQ \Right }$		
				}
			\end{tabular}		
		\end{minipage}
	}		
	\caption{System $\labIKp$}
	\label{fig:labIKp}
\end{figure}

@@

Structural proof theoretic accounts of intuitionistic modal logic can adopt the paradigm of \emph{labelled deduction} in the form of labelled natural deduction and labelled sequent systems~\cite{Simpson}, or the one of \emph{unlabelled deduction} in the form of sequent~\cite{Bierman} or nested sequent systems~\cite{Strassburger} (for a survey see~\cite[Chap.~3]{Marin}).

Simpson's proposed labelled sequent system for intuitionistic modal logic is representing explicitly only the accessibility relation $\rel$ in the syntax.
%
Echoing the definition of bi-relational structures, we consider another extension of labelled deduction to the intuitionistic setting. 
%
The idea is to use two sorts of relational atoms, one for the modal relation $\rel$ and another one for the intuitionistic relation $\leq$ as proposed by Maffezioli, Naibo and Negri in~\cite{Maffezioli}. 
%

\begin{definition}
	A two-sided intuitionistic \emph{labelled sequent} is of the form $\B, \Left \SEQ \Right$ where $\B$ denotes a set of relational atoms $x \rel y$ and preorder atoms $x \le y$, and $\Left$ and $\Right$ are multi-sets of labelled formulas $\labels{x}{A}$ (for $x$ and $y$ taken from the set of labels and $A$ an intuitionistic modal formula).
\end{definition}

@@

%
We obtain a proof system $\labIKp$, displayed on Figure~\ref{fig:labIKp}, for intuitionistic modal logic in this formalism. 
%
Most rules are similar to the ones of Simpson~\cite{Simpson}, but some rules are even more explicitly in correspondence with the semantics by using the preorder atoms. 
%
In particular, the rules introducing the $\BOX$-operator correspond to the definition $(\ast)$.
%
Furthermore, our system has to incorporate the two semantic conditions ($F_1$) and ($F_2$) into the deductive rules $\rn{F_1}$ and $\rn{F_2}$, and the rules $\rn{refl}$ and $\rn{trans}$ are also necessary to ensure that the preorder atoms do behave as a preorder relation on labels.
%
%As we mentioned, we obtain a proof system $\labIKp$ which allows us to give an extension of labelled deduction to the intuitionistic world and then we prove the next theorem:

@@

where $A_{1}, . . . , A_{n}$ are the formulas that would occur on the left of the turnstile if there was a turnstile, $B_{1}, . . . , B_{k}$ are the formulas that would occur on the right of the turnstile if there was a turnstile, and $\Cx[1], . . . , \Cx[m]$ are nested sequents. We use the $\bullet$ and $\circ$ as polarities indication for formulas.

@@

In the intuitionistic setting, the validity of a modal formula has to be defined using both the $\rel$ and the $\le$ relation as:
$\force{x}{\BOX A}$ iff for all $y$ and $z$ s.t. $\futs xy$ and $\accs yz$, $\force{z}{A}$. Our system reflects exactly this definition in the rules introducing the $\BOX$-operator:

@@

In~\cite{Simpson}, Simpson extends his basic sequent system for $\IK$ to the geometric axiom family. 
%
For example, you can add the following rule:
$$\scalebox{.9}{$\vlinf{\boxbrn{4}}{\text{\footnotesize $u'$ fresh}}{\B, w \rel v, v \rel u, \Left \SEQ \Right}{\B, w \rel v, v \rel u, w \rel u, \Left \SEQ \Right}$}$$
to it and obtain a sound and complete system wrt.~$\IK$ plus the axiom
$\ax{4}\colon (\DIA\DIA A \IMP \DIA A) \AND (\BOX A \IMP \BOX\BOX A)$, that is, wrt.~to all frames in which $\rel$ is transitive.

In~\cite{Plotkin}, Plotkin and Stirling give a more general correspondence result than Theorem~\ref{thm:plotkin}, that is, for intuitionistic modal logic extended with a family of axioms wrt.~some classes of bi-relational frames.
%
For example, the frames that validate the axiom $\rn{4}_\rn\DIA \colon \DIA\DIA A \IMP \DIA A$ are exactly the ones satisfying the condition:
%\begin{center}
($\diabrn{4}$) if $w \rel v$ and $v \rel u$, there exists a $u'\in W$ s.t.~$u \le u'$ and $wRu'$.
%\end{center}

Incorporating the preorder symbol into the syntax of our sequents allows us to also obtain a sound and complete proof system for the intuitionistic modal logic extended with axiom $\rn{4}_\rn\DIA$, by designing the following rule:
$$\scalebox{.9}{$\vlinf{\diabrn{4}}{\text{\footnotesize $u'$ fresh}}{\B, w \rel v, v \rel u, \Left \SEQ \Right}{\B, w \rel v, v \rel u, u \le u', w \rel u' , \Left \SEQ \Right}$}$$

Therefore, we decompose further the formalism of labelled sequents and extend the reach of labelled deduction to the logics studied in~\cite{Plotkin}.
%
These systems enjoy cut-elimination via usual arguments, the generality of the result is subject of ongoing study.

>>>>

  \begin{abstract}
  %We present a nested sequent system for intuitionistic modal logic in which we capture explicitly two relations: one for the accessibility relation associated with the Kripke semantics for normal modal logics (represented with one bracket in a formula or context) and one for the preorder relation associated with the Kripke semantics for intuitionistic logic (represented with two brackets in a formula or context).
  %
  We present a labelled sequent system and a nested sequent system for intuitionistic modal logics equipped with two relation symbols, one for the accesibility relation associated with the Kripke semantics for modal logics and one for the preorder relation associated with the Kripke semantics for intuitionistic logic. Both systems are in close correspondence with birelational Kripke semantics for intuitionistic modal logic.
  \end{abstract}

  \begin{keyword}
  Nested sequents, Labelled sequents, Intuitionistic modal logic, Proof theory.
  \end{keyword}
 \end{frontmatter}

@@

\section{Introduction}

\todo{sell the systems! more emphasis in the nested sequent system?}

Structural proof theoretic accounts of intuitionistic modal logic can adopt the paradigm of \emph{labelled deduction} in the form of labelled natural deduction and labelled sequent systems~\cite{simpson1994}, or the one of \emph{unlabelled deduction} in the form of sequent~\cite{bierman2000} or nested sequent systems~\cite{strassburger2013} (for a survey see~\cite[Chap.~3]{marin2018}). In this article, we present a labelled sequent system and a nested sequent system for intuitionistic modal logics capturing the accesibility relation for modal logics and the intuitionistic relation $\le$. In the labelled calculus we made explicity both relations with $\rel$ and $\le$ into the system. In the nested sequent system, we represent this relations with one bracket in a sequent (for the accesibility relation) and a with two brackets in a sequent (preorder relation).

\section{Intuitionistic modal logics}
The language of {intuitionisitic modal logic} is the one of intuitionistic propositional logic with the modal operators $\BOX$ and $\DIA$, standing most generally for \emph{necessity} and \emph{possibility}.
%
Starting with a set $\mathcal{A}$ of atomic propositions, denoted $\fm a$, modal formulas are constructed from the grammar:
%
$$
\fm A \coloncolonequals
\fm a \mid \fm{A \AND A} \mid \fm{A \OR A} \mid \fm{\TOP} \mid \fm{\BOT} \mid \fm{A \IMP A} \mid \fm{\BOX A} \mid \fm{\DIA A}
$$
%
The axiomatisation that is now generally accepted as intuitionistic modal logic $\IK$ was given by Plotkin and Stirling~\cite{plotkin1986} and is equivalent to the one proposed by Fischer-Servi~\cite{fischer1984}.
%, and by Ewald~\cite{Ewald} in the case of intuitionistic tense logic. 
%
It is obtained from intuitionistic propositional logic by adding:
\begin{itemize}
	\item the \emph{necessitation rule}: $\fm{\BOX A}$ is a theorem if $\fm{A}$ is a theorem; and
	\item the following five variants of the \emph{distributivity axiom}:
	\begin{equation*}
	\label{eq:ik}%\hskip-2em
	\begin{array}[t]{r@{\;}l@{\quad}r@{\;}l@{\quad}r@{\;}l}
	\kax[1]\colon&\fm{\BOX(A\IMP B)\IMP(\BOX A\IMP\BOX B)}
	&
	\kax[3]\colon&\fm{\DIA(A\OR B)\IMP(\DIA A\OR\DIA B)}
	&
	\kax[5]\colon&\fm{\DIA\BOT\IMP\BOT}
	\\
	\kax[2]\colon&\fm{\BOX(A\IMP B)\IMP(\DIA A\IMP\DIA B)}
	&
	\kax[4]\colon&\fm{(\DIA A\IMP \BOX B)\IMP\BOX(A\IMP B)}\\%x[1ex]
	\end{array}
	\end{equation*}
\end{itemize}

@@

\begin{definition}
	A \emph{bi-relational frame} $\F$ is a triple $\langle W, R, \le \rangle$ 
	%	of a non-empty set of worlds $W$ equipped with two binary relations $R$ and $\le$, where $R$ being the modal \emph{accessibility relation} and $\le$ a preorder (\emph{i.e.} a reflexive and transitive relation), satisfying the following conditions:
	of a set of worlds $W$ equipped with an {accessibility relation} $\rel$ and a preorder $\le$ satisfying:
	\begin{enumerate}
		\item[($F_1$)] For $\lb u, \lb v, \lb{v'} \in W$, if $\accs uv$ and $\futs{v}{v'}$, there exists $\lb{u'}$ s.t.~$\futs{u}{u'}$ and $\accs{u'}{v'}$.

		\item[($F_2$)] For $\lb{u'}, \lb u, \lb v \in W$, if $\futs uv$, there exists $\lb{v'}$ s.t.~$\accs{u'}{v'}$ and $\futs{v}{v'}$.
	\end{enumerate}
	%	
\end{definition}

%\marianela{We recall the birelational models \cite{plotkin1986, ewald1986} for intuitionistic modal logics, which are a combination of the Kripke semantics for propositional intuitionistic logic and the one for classical modal logic. A \emph{birelational frame} $\F = \langle W, R, \le \rangle$ is a non-empty set $W$ of worlds together with two binary relations $\le, \rel \subseteq W \times W$, where $\le$ is a preorder(i.e., reflexive and transitive), such that the following two conditions hold:}

\begin{definition}
	A \emph{bi-relational model} $\M$ is a quadruple $\langle W, R,\le,V \rangle$ with $\langle W, R, \le \rangle$ a bi-relational frame and $V\colon W \to 2^\mathcal{A}$ a monotone valuation function, that is, a function mapping each world $\lb w$ to the subset of propositional atoms true at $\lb w$, additionally subject to:
	if $\futs{w}{w'}$ then $V(\lb w)\subseteq V(\lb{w'})$.
\end{definition}

We write $\lb w \Vdash \fm a$ if $\fm a \in V(\lb w)$, and by definition, we always have $\lb w \Vdash \fm{\TOP}$ and never that $\lb w \Vdash \fm{\BOT}$. 
%
Then the relation is extended to all formulas by induction, following the rules for both intuitionistic and modal Kripke models:

$\lb w \Vdash \fm{A \AND B}$ iff $\lb w \Vdash \fm A$ and $\lb w \Vdash \fm B$

$\lb w \Vdash \fm{A \OR B}$ iff $\lb w \Vdash \fm A$ or $\lb w \Vdash \fm B$

$\lb w \Vdash \fm{A \IMP B}$ iff for all $\lb{w'}$ with $\futs{w}{w'}$, if $\lb{w'} \Vdash \fm A$ then $\lb{w'} \Vdash \fm B$

$\lb w \Vdash \fm{\BOX A}$ iff for all $\lb{w'}$ and $\lb u$ with $\futs{w}{w'}$ and $\accs{w'}{u}$, $\lb u \Vdash \fm A$ %\hfill $(\ast)$

$\lb w \Vdash \fm{\DIA A}$ iff there exists a $\lb u$ such that $\accs wu$ and $\lb u \Vdash \fm A$.

\begin{definition}
	A formula $\fm A$ is \emph{satisfied} in a model $\M = \langle W, R, \le, V \rangle$, if for all $\lb w \in W$ we have $\lb w \Vdash \fm A$.
	%
	A formula $\fm A$ is \emph{valid} in a frame $\F = \langle W, R, \le \rangle$, if for all valuations $V$, $\fm A$ is satisfied in $\langle W, R, \le, V \rangle$.
\end{definition}

@@

\begin{theorem}[Fischer-Servi~\cite{fischer1984}, Plotkin and Stirling~\cite{plotkin1986}]\label{thm:plotkin}
	A formula $\fm A$ is a theorem of $\IK$ if and only if $\fm A$ is valid in every bi-relational frame.
\end{theorem}

@@

\begin{figure}%[h]
	\centering
	\small
	\fbox{
		\begin{minipage}{.95\textwidth}
			\begin{tabular}{@{\!}c@{\quad}c}
				$\vlinf{\rn{id}}{}{\B, \Left, \labels{x}{A} \SEQ \Right, \labels{x}{A} }{}$
				&
				$\vlinf{\llabrn\bot}{}{\B, \Left, \labels{x}{\BOT} \SEQ \Right}{}$
%				\quad
%				$\vlinf{\rlabrn\top}{}{\B, \Left \SEQ \Right, \labels{x}{\TOP}}{}$
				\\\\
				$\vlinf{\llabrn\AND}{}{\B,\Left, \labels{x}{A \AND B} \SEQ \Right}{\B, \Left, \labels{x}{A}, \labels{x}{B} \SEQ \Right}$
				&
				$\vliinf{\rlabrn\AND}{}{\B,\Left \SEQ \Right, \labels{x}{A \AND B}}{\B, \Left \SEQ \Right, \labels{x}{A}}{\B, \Left \SEQ \Right, \labels{x}{B}}$
				\\\\
				$\vliinf{\llabrn\OR}{}{\B, \Left, \labels{x}{A \OR B} \SEQ \Right}{\B, \Left, \labels{x}{A} \SEQ \Right}{\B, \Left, \labels{x}{B} \SEQ \Right}$
				&
				$\vlinf{\rlabrn\OR}{}{\B, \Left \SEQ \Right, \labels{x}{A \OR B}}{\B, \Left \SEQ \Right, \labels{x}{A}, \labels{x}{B}}$
				\\\\
				\multicolumn{2}{c}{
					$\vlinf{\llabrn\IMP}{\text{\scriptsize $\lb y$ fresh}}{\B, \Left \SEQ \Right, \labels{x}{A \IMP B}}{\B, \Left, \futs xy, \labels{y}{A} \SEQ \Right, \labels{y}{B}}$
				}
				\\\\
				\multicolumn{2}{c}{
					$\vliinf{\rlabrn\IMP}{}{\B, \futs xy, \Left, \labels{x}{A} \SEQ B \SEQ \Right}{\B, \futs xy, \Left \SEQ \Right, \labels{y}{A}}{\B, \futs xy, \Left, \labels{y}{B} \SEQ \Right}$ 
					% 		&$\vliinf{\rlabrn\IMP}{\text{\scriptsize $x \le y \in \B$}}{\B, \Left, \labels{x}{A \IMP B} \SEQ \Right}{\B, \Left \SEQ \Right, \labels{y}{A}}{\B, \Left, \labels{y}{B} \SEQ \Right}$
				}
				\\\\
				$\vlinf{\llabrn\BOX}{}{\B, \futs xy, \accs yz, \Left, \labels{x}{\BOX A} \SEQ \Right}{\B, \futs xy, \accs yz, \Left, \labels{x}{\BOX A}, \labels{z}{A} \SEQ \Right}$
				&
				$\vlinf{\rlabrn\BOX}{\text{\scriptsize $\lb y, \lb z$ fresh}}{\B, \Left \SEQ \Right, \labels{x}{\BOX A}}{\B, \futs xy, \accs yz, \Left \SEQ \Right, \labels{z}{A}}$
				\\\\
				$\vlinf{\llabrn\DIA}{\text{\scriptsize $\lb y$ fresh}}{\B, \Left, \labels{x}{\DIA A} \SEQ \Right}{\B, \accs xy, \Left, \labels{y}{A} \SEQ \Right}$
				&
				$\vlinf{\rlabrn\DIA}{}{\B, \accs xy, \Left \SEQ \Right, \labels{x}{\DIA A}}{\B, \accs xy, \Left \SEQ \Right, \labels{x}{\DIA A}, \labels{y}{A}}$
				\\
				\multicolumn{2}{c}{
					$\mbox{\hbox to .9\linewidth{\dotfill}}$
				}
				\\
				$\vlinf{\rn{refl}}{}{\B, \Left \SEQ \Right}{\B, \futs xx, \Left \SEQ \Right}$
				&
				$\vlinf{\rn{trans}}{}{\B, \futs xy, \futs yz, \Left \SEQ \Right}{\B, \futs xy, \futs yz, \futs xz, \Left \SEQ \Right}$
				\\\\
				\multicolumn{2}{c}{
					$\vlinf{\rn{F_1}}{\text{\scriptsize $\lb u$ fresh}}{\B, \accs xy, \futs yz, \Left \SEQ \Right}{\B, \accs xy, \futs yz, \futs xu, \accs uz, \Left \SEQ \Right}$
				}
				\\\\
				\multicolumn{2}{c}{
					$\vlinf{\rn{F_2}}{\text{\scriptsize $\lb u$ fresh}}{\B, \accs xy,\futs xz, \Left \SEQ \Right}{\B, \accs xy, \futs xz, \futs yu, \accs zu, \Left \SEQ \Right }$		
				}
			\end{tabular}		
		\end{minipage}
	}		
	\caption{System $\labIKp$}
	\label{fig:labIKp}
\end{figure}

@@

\emph{Labelled deduction} has been proposed by Gabbay~\cite{gabbay1996} in the 80s as a unifying framework throughout proof theory in order to provide proof
systems for a wide range of logics. 
%
For modal logics it can take
the form of labelled natural deduction and labelled sequent systems as used, for example, by Simpson~\cite{simpson1994}, Vigan\`o~\cite{vigano2000}, and Negri~\cite{negri2005}. 
%
Simpson's proposed labelled sequent system for intuitionistic modal logic is representing explicitly only the accessibility relation $\rel$ in the syntax.
%
Echoing the definition of bi-relational structures, we consider another extension of labelled deduction to the intuitionistic setting. 
%
The idea is to use two sorts of relational atoms, one for the modal relation $\rel$ and another one for the intuitionistic relation $\leq$ as proposed by Maffezioli, Naibo and Negri in~\cite{maffezioli2013}. 
%

\begin{definition}
	A two-sided intuitionistic \emph{labelled sequent} is of the form $\B, \Left \SEQ \Right$ where $\B$ denotes a set of relational atoms $\accs xy$ and preorder atoms $\futs xy$, and $\Left$ and $\Right$ are multi-sets of labelled formulas $\labels{x}{A}$ (for $\lb x$ and $\lb y$ taken from the set of labels and $A$ an intuitionistic modal formula).
\end{definition}

@@

%
We obtain a proof system $\labIKp$, displayed on Figure~\ref{fig:labIKp}, for intuitionistic modal logic in this formalism. 
%
Most rules are similar to the ones of Simpson~\cite{simpson1994}, but some rules are even more explicitly in correspondence with the semantics by using the preorder atoms. 
%
In particular, the rules introducing the $\BOX$-operator correspond to the definition $(\ast)$.
%
Furthermore, our system has to incorporate the two semantic conditions ($F_1$) and ($F_2$) into the deductive rules $\rn{F_1}$ and $\rn{F_2}$, and the rules $\rn{refl}$ and $\rn{trans}$ are also necessary to ensure that the preorder atoms do behave as a preorder relation on labels.
%
%As we mentioned, we obtain a proof system $\labIKp$ which allows us to give an extension of labelled deduction to the intuitionistic world and then we prove the next theorem:

@@

where $\fm{A_{1}}, . . . , \fm{A_{n}}$ are the formulas that would occur on the left of the turnstile if there was a turnstile, $\fm{B_{1}}, . . . , \fm{B_{k}}$ are the formulas that would occur on the right of the turnstile if there was a turnstile, and $\Cx[1], . . . , \Cx[m]$ are nested sequents. We use the $\bullet$ and $\circ$ as polarities indication for formulas.

@@

In the intuitionistic setting, the validity of a modal formula has to be defined using both the $\rel$ and the $\le$ relation as:
$\force{\lb x}{\fm{\BOX A}}$ iff for all $\lb y$ and $\lb z$ s.t. $\futs xy$ and $\accs yz$, $\force{\lb z}{\fm{A}}$. Our system reflects exactly this definition in the rules introducing the $\BOX$-operator:

@@

In~\cite{Simpson}, Simpson extends his basic sequent system for $\IK$ to the geometric axiom family. 
%
For example, you can add the following rule:
$$\scalebox{.9}{$\vlinf{\boxbrn{4}}{\text{\footnotesize $\lb{u'}$ fresh}}{\B, \accs wv, \accs vu, \Left \SEQ \Right}{\B, \accs wv, \accs vu, \accs wu, \Left \SEQ \Right}$}$$
to it and obtain a sound and complete system wrt.~$\IK$ plus the axiom
$\ax{4}\colon \fm{(\DIA\DIA A \IMP \DIA A) \AND (\BOX A \IMP \BOX\BOX A)}$, that is, wrt.~to all frames in which $\rel$ is transitive.

In~\cite{plotkin1986}, Plotkin and Stirling give a more general correspondence result than Theorem~\ref{thm:plotkin}, that is, for intuitionistic modal logic extended with a family of axioms wrt.~some classes of bi-relational frames.
%
For example, the frames that validate the axiom $\rn{4}_\rn\DIA \colon \fm{\DIA\DIA A \IMP \DIA A}$ are exactly the ones satisfying the condition:
%\begin{center}
($\diabrn{4}$) if $\accs wv$ and $\accs vu$, there exists a $\lb{u'}\in W$ s.t.~$\futs{u}{u'}$ and $\accs{w}{u'}$.
%\end{center}

Incorporating the preorder symbol into the syntax of our sequents allows us to also obtain a sound and complete proof system for the intuitionistic modal logic extended with axiom $\rn{4}_\rn\DIA$, by designing the following rule:
$$\scalebox{.9}{$\vlinf{\diabrn{4}}{\text{\footnotesize $\lb{u'}$ fresh}}{\B, \accs wv, \accs vu, \Left \SEQ \Right}{\B, \accs wv, \accs vu, \futs{u}{u'}, \accs{w}{u'}, \Left \SEQ \Right}$}$$

Therefore, we decompose further the formalism of labelled sequents and extend the reach of labelled deduction to the logics studied in~\cite{plotkin1986}.
%
These systems enjoy cut-elimination via usual arguments, the generality of the result is subject of ongoing study.
